Sentence Generation as Planning with Probabilistic LTAG\\

### 2010-05-21 trunk/Papers/pcrisp-10/conclusion.tex
einmal ganz drueber

We have described PCRISP, an approach to integrated sentence generation which can compute the best derivation according to a probabilistic TAG grammar.  This brings two strands of research -- statistical generation and integrated sentence planning and realization -- together for the first time.  Our generation algorithm operates by converting the generation problem into a metric planning problem and solving it with an off-the-shelf planner.  An evaluation on the WSJ corpus revealed that while PCRISP is excellent at computing the preferred derivation on seen data, it is still seriously hampered by data sparseness issues on unseen data.  Because the size of the planning problem is quadratic in the number of lexicalized trees in the grammar, current planning algorithms are also too slow to be used for longer sentences.

An obvious issue for future research would be to apply improved smoothing techniques to deal with the data sparseness.  Planning runtimes should be improved by further tweaking the exact planning problems we generate, and will benefit from any future improvements in metric planning.  It is interesting to note that the extensions we made to CRISP to accommodate statistical generation here are compatible with recent work in which CRISP is applied to situated generation \cite{garoufikoller2010}; we expect that this will be true for other future extensions to CRISP as well.  Finally, we have only evaluated PCRISP on a surface realization problem in this paper.  It would be interesting to carry out an extrinsic, task-based evaluation of PCRISP that also addresses sentence planning.

\paragraph{Acknowledgments.} We are grateful to Owen Rambow for providing us with the Chen WSJ-TAG corpus and to Malte Helmert and Silvia Richter for their help with running LAMA, another metric planner with which we experimented. We thank Owen and Konstantina Garoufi for helpful discussions, and our reviewers for their insightful comments.

>>>>

We have described PCRISP, an approach to integrated sentence generation which can compute the best derivation according to a probabilistic TAG grammar.  This brings two strands of research -- statistical generation and integrated sentence planning and realization -- together for the first time.  Our generation algorithm operates by converting the generation problem into a metric planning problem and solving it with an off-the-shelf planner.  An evaluation on the WSJ corpus reveals that \pcrisp, like PTAG in general, is susceptible to data sparseness problems.  Because the size of the planning problem is quadratic in the number of lexicalized trees in the grammar, current planning algorithms are also too slow to be used for longer sentences.

An obvious issue for future research is to apply improved smoothing techniques to deal with the data sparseness.  Planning runtimes should be improved by further tweaking the exact planning problems we generate, and will benefit from any future improvements in metric planning.  It is interesting to note that the extensions we made to CRISP to accommodate statistical generation here are compatible with recent work in which CRISP is applied to situated generation \cite{garoufikoller2010}; we expect that this will be true for other future extensions to CRISP as well.  Finally, we have only evaluated PCRISP on a surface realization problem in this paper.  It would be interesting to carry out an extrinsic, task-based evaluation of PCRISP that also addresses sentence planning.

\paragraph{Acknowledgments.} We are grateful to Owen Rambow for providing us with the Chen WSJ-TAG corpus and to Malte Helmert and Silvia Richter for their help with running LAMA, another metric planner with which we experimented. We thank Konstantina Garoufi and Owen Rambow for helpful discussions, and our reviewers for their insightful comments.

### 2010-05-21 trunk/Papers/pcrisp-10/crisp.tex
einmal ganz drueber

\section{Sentence Generation as Planning}
\label{sec:crisp}
In this section we review the original non-statistical {\sc crisp} system \cite{kollerstone2007}. 
Following {\sc spud} \cite{stonedoran1997}, {\sc crisp} is based on a declarative description of the sentence generation problem using {\sc tag}. Given a knowledge base, a communicative goal and a grammar, we require to find a grammatical {\sc tag} derivation that is consistent with this knowledge base and satisfies a communicative goal. A number of semantic and pragmatic constraints that must be satisfied by the solution can be added, for instance to enforce generation of unambiguous referring expressions. \newcite{kollerstone2007} describe how to encode this problem into an AI planning problem, that can be solved efficiently by off-the-shelf planners. We describe the general mechanism in the following section and then review the encoding into planning in section \ref{ssec:crispdomain}. 
\begin{figure*}[th]
\begin{center}
\includegraphics[width=.8\textwidth]{figures/grammar.pdf}
\caption{\label{fig:grammar}(a) An example grammar with semantic content. (b) A derivation for ``the cat eats the raw fish''. }
\end{center}
\end{figure*}

\subsection{Sentence Generation in CRISP}
Like {\sc spud}, {\sc crisp} uses an {\sc ltag} in which elementary trees are assigned semantic content. Each node in a {\sc crisp} elementary tree is associated with a semantic role. Semantic content is expressed as a set of literals, encoding relations between these roles. All nodes that dominate the lexical anchor are assigned the role `self', which intuitively corresponds to the event or individual described by this tree. Figure \ref{fig:grammar}(a) shows an example grammar of this type.

In a derivation we may only include elementary trees whose semantic content has an instantiation in the knowledge base. For each substitution and adjunction, the semantic role associated with the role of the target node is unified with the `self' role of the child tree. For example, given the knowledge base 
{\it \{cat($i_2$), fish($i_3$), raw($i_3$), fish($i_4$), eats($i_1$,$i_2$,$i_3$)\}} and the grammar in \ref{fig:grammar}(a), {\sc crisp} could produce the derivation in \ref{fig:grammar}(b). Notice that {\sc crisp} can generate the unambiguous referring expression `the raw fish' for $i_3$, to distinguish it from $i_4$. For details on  referring expression generation in {\sc crisp} see \newcite{garoufikoller2010}.

\subsection{CRISP Planning Domains} 
\label{ssec:crispdomain}
Before we describe {\sc crisp}'s encoding of sentence generation as planning, we briefly review AI planning in general. 
A planning state is a conjunction of first order literals describing relations between some individuals. A planning problem consist of an initial state, a set of goal states and a set of planning operators that describe possible state transitions.  A solution to the problem is any sequence of actions (instantiated operators) that leads from the initial state to one of the goal states. Planning problems can be solved efficiently by general purpose planning systems such as 
{\sc ff} \cite{hoffmannnebel2001}.  
% Planning problems can be written in a standard notation ({\sc pddl}\footnote{Planning Domain Description Language}, \newcite{gerevinietal2009}).

In {\sc crisp} planning states correspond to partial {\sc tag} derivations and record open substitution and adjunction sites, semantic individuals associated with them, and parts of the communicative goal that have not yet been expressed.  The initial state also encodes the knowledge base and the communicative goal. Each planning operator contributes a new elementary tree to the derivation and at the same time satisfies part of the communicative goal, as described in the previous section. In a goal state there are no open substitution sites left and all literals in the communicative goal have been expressed. 
\begin{figure}[t]
\begin{center}
\planaction{\bf subst-t28-eats-S(u,~x1,~x2,~x3)}{referent(u,~x1),\\ subst(S,~u), eats(x1,~x2,~x3)}{
$\lnot$needtoexpr(pred-eats,~x1,~x2,~x3),\\ $\lnot$subst(S,~u),\\
subst(NP,~subj), subst(NP,~obj),\\ referent(subj,~x2), referent(obj,~x3)\\
adj(VP, u), adj(V, u), adj(S, u)}\\\smallskip

@@

\planaction{\bf subst-t5-raw-N(u,~x1)}{referent(u,~x1),\\ adj(N,~u), raw(x1)}{
$\lnot$needtoexpr(pred-raw,~x1),\\ $\lnot$adj(N,~u)}\\\smallskip
\end{center}
\caption{\label{fig:crisp-operators} {\sc crisp} operators for some of the elementary trees in figure \ref{fig:grammar}.}
\end{figure}

Figure \ref{fig:crisp-operators} shows planning operators for part of the grammar in figure \ref{fig:grammar}(a).
The preconditions of the operators require that a suitable open substitution node (i.e. of the correct category) or internal node for adjunction exists in the partial derivation. In the operator effect, open substitution nodes are closed and new identifiers are created for each substitution node and internal node in the new tree. Given the knowledge base from above, a plan corresponding to the derivation in figure \ref{fig:grammar} would be $\langle${\it S-eats-1(root, i1, i2, i3), NP-cat-2(subj-1, i2), NP-fish-3(obj-1, i3), N-raw-4(obj-1, i3)}$\rangle$

\subsection{CRISP and Large Grammars}
Using {\sc crisp} with large, treebank induced grammars can lead to overgeneration, 
%Such trees contain many internal nodes that are open for adjunction but will not necessarily be filled in a derivation for a given input communicative goal. Attaching to the wrong node can produce invalid or dispreferred word order. 
due to the high lexical ambiguity of possible, structurally different trees (supertags) for a given word. We illustrate this problem in figure \ref{fig:overgen}. Assuming a grammar that includes trees for both right adjoining ($t13$) and left adjoining PPs ($t252$), both derivations (a) and (b) are {\it grammatical} derivations that satisfy the same communicative goal. However, most readers disprefer the reading in (b). 
Clearly, to use {\sc crisp} with such a grammar we need a method of distinguishing good derivations from bad ones. In section \ref{sec:experiments} we provide empirical evidence for this problem and show how it is solved by the approach we describe in the next section. 
\begin{figure}
\begin{center}
\includegraphics[width=.3\textwidth]{figures/overgen.pdf}
\caption{\label{fig:overgen} Two derivations with a large grammar, that satisfy the same communicative goal. Sentence (b) is dispreferred by most readers.}
\end{center}
\end{figure}

>>>>

\section{Sentence Generation as Planning}
\label{sec:crisp}
In this section we review the original non-statistical {\sc crisp} system \cite{kollerstone2007}. 
Following {\sc spud} \cite{Stone2003a}, {\sc crisp} is based on a declarative description of the sentence generation problem using {\sc tag}. Given a knowledge base, a communicative goal and a grammar, we require to find a grammatical {\sc tag} derivation that is consistent with this knowledge base and satisfies a communicative goal. A number of semantic and pragmatic constraints that must be satisfied by the solution can be added, for instance to enforce generation of unambiguous referring expressions. \newcite{kollerstone2007} describe how to encode this problem into an AI planning problem which can be solved efficiently by off-the-shelf planners. We describe the general mechanism in the following section and then review the encoding into planning in section \ref{ssec:crispdomain}. 
\begin{figure*}[th]
\begin{center}
\includegraphics[width=.8\textwidth]{figures/grammar.pdf}
\caption{\label{fig:grammar}(a) An example grammar with semantic content. (b) A derivation for ``the cat eats the raw fish''. }
\end{center}
\end{figure*}

\subsection{Sentence Generation in CRISP}
Like {\sc spud}, {\sc crisp} uses an {\sc ltag} in which elementary trees are assigned semantic content. Each node in a {\sc crisp} elementary tree is associated with a semantic role. Semantic content is expressed as a set of literals, encoding relations between these roles. All nodes that dominate the lexical anchor are assigned the role `self', which intuitively corresponds to the event or individual described by this tree. Fig.~\ref{fig:grammar}(a) shows an example grammar of this type.

In a derivation we may only include elementary trees whose semantic content has an instantiation in the knowledge base. For each substitution and adjunction, the semantic role associated with the role of the target node is unified with the `self' role of the child tree. For example, given the knowledge base $\{\f{cat}(c), \f{fish}(f_1), \f{raw}(f_1), \f{fish}(f_2), \f{eats}(e,c,f_1)\}$ and the grammar in \ref{fig:grammar}(a), {\sc crisp} could produce the derivation in \ref{fig:grammar}(b). Notice that {\sc crisp} can generate the unambiguous referring expression `the raw fish' for $f_1$, to distinguish it from $f_2$.

\subsection{CRISP Planning Domains} 
\label{ssec:crispdomain}
Before we describe {\sc crisp}'s encoding of sentence generation as planning, we briefly review AI planning in general.  A planning state is a conjunction of first order literals describing relations between some individuals. A planning problem consist of an initial state, a set of goal states and a set of planning operators that describe possible state transitions.  A \emph{plan} is any sequence of actions (instantiated operators) that leads from the initial state to one of the goal states. Planning problems can be solved efficiently by general purpose planning systems such as {\sc ff} \cite{hoffmannnebel2001}.

In {\sc crisp}, planning states correspond to partial {\sc tag} derivations and record open substitution and adjunction sites, semantic individuals associated with them, and parts of the communicative goal that have not yet been expressed.  The initial state also encodes the knowledge base and the communicative goal. Each planning operator contributes a new elementary tree to the derivation and at the same time can satisfy part of the communicative goal, as described in the previous section. In a goal state there are no open substitution sites left and all literals in the communicative goal have been expressed. 
\begin{figure}[t]
\begin{center}
\planaction{\bf subst-t28-eats-S(u,~x1,~x2,~x3)}{referent(u,~x1),\\ subst(S,~u), eats(x1,~x2,~x3)}{
$\lnot$needtoexpr(pred-eats,~x1,~x2,~x3),\\ $\lnot$subst(S,~u),\\
subst(NP,~subj), subst(NP,~obj),\\ referent(subj,~x2), referent(obj,~x3)\\
adj(VP, u), adj(V, u), adj(S, u)}\\\smallskip

@@

\planaction{\bf adj-t5-raw-N(u,~x1)}{referent(u,~x1),\\ adj(N,~u), raw(x1)}{
$\lnot$needtoexpr(pred-raw,~x1),\\ $\lnot$adj(N,~u)}\\\smallskip
\end{center}
\caption{\label{fig:crisp-operators} {\sc crisp} operators for some of the elementary trees in Fig.~\ref{fig:grammar}.}
\end{figure}

Fig.~\ref{fig:crisp-operators} shows planning operators for part of the grammar in Fig.~\ref{fig:grammar}(a). The operators are simplified for lack of space, and in particular we do not show the preconditions and effects that enforce uniqueness of referring expressions; see \newcite{kollerstone2007} for details.  The preconditions of the operators require that a suitable open substitution node (i.e. of the correct category) or internal node for adjunction exists in the partial derivation. In the operator effect, open substitution nodes are closed and new identifiers are created for each substitution node and internal node in the new tree. Given the knowledge base from above, a plan corresponding to the derivation in Fig.~\ref{fig:grammar} would be $\f{subst-t28-eats-S-eats}(\f{root}, e, c, f_1)$; $\f{subst-t3-cat-NP}(\f{subj},c)$; $\f{subst-t3-fish-NP}(\f{obj},f_1)$; $\f{adj-t5-raw-N}(\f{obj}, f_1)$. This plan can be automatically decoded into a derivation tree for Fig.~\ref{fig:grammar}b.

\subsection{CRISP and Large Grammars}

\newcite{KolHof10} report on experiments that show CRISP can generate sentences with the large-scale XTAG grammar \cite{xtag2001} quite efficiently.  However, because CRISP has no notion of how ``good'' a generated sentence is compared to other grammatical alternatives, it will sometimes compute dispreferred sentences with large grammars.  This is especially true for treebank-induced grammars, which tend to overgenerate and rely on statistical methods to rank good sentences highly.  Fig.~\ref{fig:overgen} illustrates this problem.  Assuming a (treebank) grammar that includes trees for both right adjoining (t13) and left adjoining PPs (t252), both derivations (a) and (b) are {\it grammatical} derivations that satisfy the same communicative goal. However, most readers disprefer the reading in (b).  Clearly, to use {\sc crisp} with such a grammar we need a method of distinguishing good derivations from bad ones.

\begin{figure}
\begin{center}
\includegraphics[width=.3\textwidth]{figures/overgen.pdf}
\caption{\label{fig:overgen} Two derivations with a large grammar, that satisfy the same communicative goal. Sentence (b) is dispreferred by most readers.}
\end{center}
\end{figure}

%%% Local Variables:  %%% mode: latex %%% TeX-master: "pcrisp-10" %%% End:

### 2010-05-21 trunk/Papers/pcrisp-10/experiments.tex
einmal ganz drueber

\section{Experiments}
\label{sec:experiments}
In this section we perform experiments with the statistical generation system described in the previous section.  We are interested in the output quality and runtime behavior of different combinations of planning systems and probability models. Like {\sc crisp}, {\sc pcrisp} is a sentence generation system and can for instance generate referring expressions. Unfortunately, since we do not have an evaluation dataset for full sentence generation, we only evaluate the realization abilities of our system by comparing generated sentences against a gold standard testset. The next section describes our evaluation data and section \ref{ssec:eval-measures} describes the evaluation measures used.

\subsection{Evaluation Data}
For our experiments we use a grammar and treebank that was automatically extracted from the Wall Street Journal using the algorithm described by \newcite{chenschanker2004}.

This algorithm outputs a grammar that allows multiple adjunctions at the same node. For such a grammar {\sc ptag} is not a suitable probability model. Models that can deal with multiple adjunction are discussed by \newcite{nasrrambow2006}, but employing them would require non-trivial modifications to our encoding as a metric planning problem. We therefore preprocess the treebank by `linearizing' multiple adjunctions. Furthermore, we attach prepositions to the trees they substitute in, to increase the expressiveness of the bilexical probability distribution. For details on these preprocessing steps see \cite{bauer2009}.

%{\bf Adding semantics to elementary trees:}
We then automatically create semantic content for all lexicalized elementary trees by assigning a single semantic predicate to each tree in the treebank, using the lexical anchor as a name and setting the arity to the number of substitution nodes in the tree plus one for the `self' role.  We calculate the role associated with each node in each tree by assigning role-names to each substitution node (`self' is assigned to the lexical anchor) and then percolating the roles up the tree, giving preference to the self role.

We estimate our probability models on a training portion of the dataset (96\%) using maximum likelihood estimation.
For the testing set we selected from the remaining 4\% all sentences that were shorter than 16 words (416 sentences). For longer sentences the search time often exceeded 5 minutes.

%{\bf Creating realization problems:} 
For this testing set we compute semantic representation for each sentence, by traversing its derivation tree top-down and collects a conjunction of terms, by instantiating the automatically created semantic representations. Our algorithm introduces a new individual for each substitution node marked with role $r$ and equates it to the `self' role of any tree that adjoins or substitutes into a node marked with $r$.

\subsection{Evaluation Measures}
\label{ssec:eval-measures}
To evaluate the output quality of our statistical generation system we compare the system output $O$ against the reference sentence $R$in the treebank, from which the input representation was generated. We adopt the {\bf generation tree accuracy} measure discussed by \cite{bangalorerambowwhittaker2001}.
This measure is computed by first creating a list of all 'treelets' from the reference derivation tree $D$. A 'treelet' is a subtree, consisting only of a node and its direct children. For each treelet we calculate edit distance, sum the distances over all treelets and then divide by the total length of the reference string.
$$1 - \frac{\sum\limits_{d \in treelets(D)} editDist(O|_d, R|_d)}{\textit{length}(R)},$$
where $D$ is the reference derivation tree and $S|_d$ are the tokens associated with the nodes of treelet $d$ in the order they appear in $S$ (if at all). 
Edit distance is modified to treat insertion-deletion pairs as single movement errors. 
This metric account for the observation that errors that change word order on a local level (i.e. within a treelet) are less severe than errors that swap words between different treelets. Intuitively errors of the latter type violate the continuity of sub-phrases. Because of this property we prefer generation tree accuracy over {\sc blue}, which is overly sensitive to movement errors on an individual sentence level.

%\newcite{belzreiter2006} empirically investigate the usefulness of {\sc bleu} and {\sc nist} for evaluating NLG systems. They find that both metrics correlate with human judgments of output quality on a corpus level, but require a large number of different references sentences to be reliable. Multiple references are necessary because {\sc blue} favors sentences with the same word order as the reference. In NLG, more than in Machine Translation, word order of output sentences can vary.

%{\it String Accuracy} is defined as 
%$$1 - \frac{\textit{editDist(O,R)}}{\textit{length}(R)},$$ where $O$ is the surface system output, $R$ is the reference sentence and $editDist(A,B)$ is the minimal edit distance between the token strings $A$ and $B$ defined as the minimal number of substitutions, insertions and deletions needed to transform $A$ into $B$.

%Like {\sc bleu}, string accuracy often underestimates generation performance if the system output differs from the reference only in word order. 
%For example if a word occurs in the beginning of the string instead of the end, blocked by a number of correct words, edit distance will count two errors, one for removing the word in the beginning and one for adding it in the end. In this case it would be more reasonable to count a single movement error.
%The revised {\it generation string accuracy} does this in a post-processing step: Each insertion-deletion pair sharing the same token, but in different places is counted as a single movement error. 
%\newcite{bangalorerambowwhittaker2001}  refine these string based measures by referring to structural information from the reference derivation tree.

%Tree accuracy measure only the quality of the surface string. To compare the syntactic structure of the output with the reference we propose {\it derivation tree accuracy}, the ratio of derivation-treelets (nodes with their direct children in their specific order) that occurs in both the system output derivation and the reference derivation. 
%$$ \frac{|treelets(D_O) \cap treelets(D_R)|}{\textit{length}(D_R)}$$
% Notice that measures used in statistical parsing, e.g. bracketing and labeling accuracy, cannot be used to evaluate generation output because they can only measure similarity between two different syntactic structures for the same surface string.

\subsection{Planning Systems}
The planning system we used was a reimplementation of {\sc FF} \cite{hoffmannnebel2001}, a heuristic search planner. We extended this planner to take action costs into account when it calculates heuristic values for successor states. We also experimented with the greedy search heuristic used by {\sc spud} \cite{stonedoran1997}.

\subsection{Results}
Table \ref{results} presents our results.

The first two columns confirm our hypothesis that {\sc crisp} does not scale up to the automatically induced treebank. The greedy {\sc spud} heuristic does not generate a single correct sentence, as the combinatorics of the planning domain becomes too large with the treebank induced grammar. 
With the standard {\sc ptag} definition our system manages to generate only very few sentences due to data sparseness. However these sentences are almost identical to the ones found in the test set. Both generalized models ease the data sparseness problem, but output quality drops rapidly. For the unlexicalized model the search problem becomes very hard due to the large planning domain, so that 22\% of the generation problems require more than 5 minutes runtime. As expected the smoothed model offers a compromise between efficiency, output quality and coverage 
\footnote{In our experiments we set the weight between level 1 and 2 to 0.9 and did not consider level 3, after empirically testing different settings. We set a probability threshold to include operators into the planning domain to $10^{-4}$}.

\begin{table}
    \caption{\label{results} Evaluation results.}
    \begin{center}
    \begin{tabular}{|l|l|r|r|r|}
    \multicolumn{5}{c}{\sc crisp}\\ 
    \hline
     {\bf Planner }  & {\bf acc}& {\bf success} & {\bf fail} & {\bf timeout} \\ \hline 
     greedy         & ??? & 0\% & 100\% & 0\%  \\ \hline 
     FF         &  0.66 & 88\% & 0\% & 12\% \\ \hline 
    \multicolumn{5}{c}{\bf planning with {\sc ptag}}\\\hline
    {\bf Planner }  & {\bf acc}& {\bf success} & {\bf fail} & {\bf timeout} \\ \hline 
     FF + costs &  0.99 &  97\% & 0\% & 3\% \\ \hline
     FF         &  0.99 & 71\% & 0\% &  29\% \\ \hline
    \multicolumn{5}{c}{\bf planning with {\sc ptag} (unlexicalized)}\\\hline
    {\bf Planner }  & {\bf acc}& {\bf success} & {\bf fail} & {\bf timeout} \\ \hline 
     FF + costs & 0.74 & 62\% & 16\% & 22\%\\ \hline
     FF & 0.76 & 62\% & 16\% &  22\%\\ \hline
    \multicolumn{5}{c}{\bf planning with {\sc ptag} (smoothed)}\\\hline
    {\bf Planner }  & {\bf acc}& {\bf success} & {\bf fail} & {\bf timeout} \\ \hline 
     FF + costs & 0.78 & 19\% & 74\% & 7\% \\ \hline
     FF & 0.77 & 19\% & 74\% & 7\% \\ \hline
    \end{tabular}
    \end{center}
\end{table}

>>>>

\section{Evaluation}
\label{sec:experiments}

We now report on some experiments with \pcrisp. We are interested in the output quality and runtime behavior of different combinations of planning systems and probability models. Like \crisp, \pcrisp\ is an integrated sentence generation system capable of generating referring expressions during surface realization.  However, for lack of an appropriate evaluation dataset for full sentence generation, we only evaluate \pcrisp\ on a realization task here.  Further experiments and more details can be found in \newcite{bauer2009}.

\subsection{Evaluation Data}
For our experiments we use an \ltag\ grammar and treebank that was automatically extracted from the Wall Street Journal using the algorithm described by \newcite{chenschanker2004}.

This algorithm outputs a grammar that allows multiple adjunctions at the same node. For such a grammar {\sc ptag} is not a suitable probability model. Models that can deal with multiple adjunction are discussed by \newcite{nasrrambow2006}, but employing them would require non-trivial modifications to our encoding as a metric planning problem. We therefore preprocess the treebank by linearizing multiple adjunctions. Furthermore, we reattach prepositions to the trees they substitute in, to increase the expressiveness of the bilexical probability distribution.

We then automatically create semantic content for all lexicalized elementary trees by assigning a single semantic literal to each tree in the treebank, using the lexical anchor as the predicate symbol and variables for each substitution node and the `self' role as arguments.  We calculate the role associated with each node in each tree by assigning role names to each substitution node (`self' is assigned to the lexical anchor) and then percolating the roles up the tree, giving preference to the `self' role.

We estimate our probability models on section 1 to 23 of the converted WSJ using maximum likelihood estimation. We use section 0 as a testing set. However, since the number of \pcrisp\ operators grows quadratically with the grammar size, generating long sentences requires too much time to run batch experiments.  We therefore restricted our evaluation to the 416 sentences in Section 0 that are shorter than 16 words.

For this testing set we automatically create semantic representation for each sentence, by instantiating the semantic content of each elementary used in its derivation. We use these representations as input for our system and compare the system output against the original sentence.

\subsection{Generation tree accuracy}
\label{ssec:eval-measures}
To evaluate the output quality of our statistical generation system we compare the system output $O$ against the reference sentence $R$ in the treebank from which the input representation was generated. We adopt the {\bf generation tree accuracy} (GTA) measure discussed by \cite{bangalorerambowwhittaker2001}.
This measure is computed by first creating a list of all 'treelets' from the reference derivation tree $D$. A `treelet' is a subtree consisting only of a node and its direct children. For each treelet we calculate the edit distance, sum the distances over all treelets and then divide by the total length of the reference string:
$$1 - \frac{\sum\limits_{d \in treelets(D)} editDist(O|_d, R|_d)}{\textit{length}(R)},$$
where $D$ is the reference derivation tree and $S|_d$ are the tokens associated with the nodes of treelet $d$ in the order they appear in $S$ (if at all).  Edit distance is modified to treat insertion-deletion pairs as single movement errors.  Compared to a purely string-based metric like {\sc bleu}, GTA penalizes swapped words less harshly if they can be explained by local tree movements.

\subsection{Results}

Table \ref{results} presents the results of the experiment for five different generation systems.  We compare three variants of \pcrisp: the fully lexicalized PTAG model (``PTAG''), the fully unlexicalized model (``unlexicalized''), and a linear interpolation model in which we (manually) set the weight of level 2 to 0.9 \todo{stimmt das?} and the weight of level 3 to 0 (``interpolation'').  We also list results for the non-statistical \crisp\ system of  \newcite{kollerstone2007} (``\crisp'') and the greedy search heuristic used by \spud.  All these systems are based on a reimplementation of the FF planner \cite{hoffmannnebel2001}, to which we added a search heuristic that takes action costs into account for the \pcrisp\ systems.

For each system, we determine the proportion of sentences in the test set for which the system produced an output (``success''), did not find a plan (``fail''), and exceeded the five-minute timeout (``timeout'').  The column ``gta'' records the mean generation tree accuracy for those sentences where the system produced an output.

The table confirms that the non-statistical \crisp\ system has considerable trouble reconstructing the original sentences from the treebank, with a mean GTA of 0.66.  This is still better than our reimplementation of \spud, which fails to recover from early mistakes of its greedy search heuristic for every single sentence in the test set.

By contrast, the fully lexicalized \pcrisp\ model achieves a much better mean GTA.  However, this comes at the cost of a very low success rate of only 10\%, reflecting a serious data sparseness problem on unseen inputs.  The data sparseness problem is reduced in the unlexicalized version of \pcrisp, but this comes at the cost of decreased accuracy and much increased runtimes.  The linear interpolation model strikes a balance between these two, by improving the success rate over the lexicalized model while sacrificing only a small amount of accuracy.  This suggests that smoothing is a promising approach to balancing coverage, efficiency, and accuracy, but clearly further experimentation is needed to substantiate this.

\begin{table}
    \begin{center}
    \begin{tabular}{|l||l||r|r|r|}
    \hline
     {\bf System }  & {\bf gta}& {\bf success} & {\bf fail} & {\bf timeout} \\ \hline 
     \spud         & n/a & 0\% & 100\% & 0\%  \\ \hline 
     \crisp         &  0.66 & 45\% & 42\% & 13\% \\ \hline
     PTAG &  0.90 &  10\% &88\% & 2\% \\ \hline
     unlexicalized & 0.74 & 62\% & 16\% & 22\%\\ \hline
     interpolation & 0.88 & 19\% & 74\% & 7\% \\ \hline
    \end{tabular}
    \end{center}
    \caption{\label{results} Results for the realization experiment.} 
\end{table}

%%% Local Variables:  %%% mode: latex %%% TeX-master: "pcrisp-10" %%% End:

### 2010-05-21 trunk/Papers/pcrisp-10/introduction.tex
einmal ganz drueber

Sentence generation systems are typically organized in a pipeline
architecture, in which the semantic input representation is first
enriched by such things as referring expressions by a \emph{sentence
  planner} and only then transformed into natural language strings by
a \emph{surface realizer} \cite{reiterdale2000}.  An alternative
approach is \emph{integrated} sentence generation, in which both steps
are performed by the same algorithm, as in the SPUD system
\cite{Stone2003a}.  An integrated algorithm can sometimes generate
better and more succinct sentences \cite{stone98textual}.  SPUD itself
gives up some of this advantage by using a greedy search heuristic
for efficiency reasons.  The CRISP system, a recent reimplementation
of SPUD using search techniques from AI planning, achieves high
efficiency without sacrificing complete search
\cite{kollerstone2007,KolHof10}.

While CRISP is efficient enough to perform well on large-scale
grammars \cite{kollerpetrick2009}, such grammars tend to offer many
different ways to express the same semantic representation.  This
makes it necessary for the generation system to be able to compute not
just grammatical sentences, but to identify which of these sentences
are the \emph{best}. This problem is exacerbated when using
treebank-derived grammars, which tend to underspecify the actual
constraints on grammaticality and instead rely on statistical
information learned from the treebank.  Indeed, there have been a
number of systems for statistical generation, which can exploit such
information to rank sentences appropriately
\cite{langkildeknight1998,whitebaldridge2003,belz2008}.  However,
these systems are restricted to performing surface realization.

In this paper, we bring these two strands of research together for the
first time.  We present the PCRISP system, which redefines the SPUD
generation problem in terms of probabilistic TAG grammars (PTAG,
\cite{resnik1992}) and then extends CRISP to solving the probabilistic
SPUD generation problem using metric planning \cite{fox2002,hoffmann2003}.  We
evaluate PCRISP on a PTAG treebank extracted from the Wall Street
Journal Corpus \cite{chenschanker2004}.  While the evaluation results
are not competitive in terms of efficiency and suffer from data
sparseness problems, we still believe our system is a first step
towards probabilistic integrated generation.

>>>>

Many sentence generation systems are organized in a pipeline
architecture, in which the input semantic representation is first
enriched, e.g.\ by referring expressions, by a \emph{sentence
  planner} and only then transformed into natural language strings by
a \emph{surface realizer} \cite{reiterdale2000}.  An alternative
approach is \emph{integrated} sentence generation, in which both steps
are performed by the same algorithm, as in the SPUD system
\cite{Stone2003a}.  An integrated algorithm can sometimes generate
better and more succinct sentences \cite{stone98textual}.  SPUD itself
gives up some of this advantage by using a greedy search heuristic
for efficiency reasons.  The CRISP system, a recent reimplementation
of SPUD using search techniques from AI planning, achieves high
efficiency without sacrificing complete search
\cite{kollerstone2007,KolHof10}.

While CRISP is efficient enough to perform well on large-scale
grammars \cite{KolHof10}, such grammars tend to offer many
different ways to express the same semantic representation.  This
makes it necessary for the generation system to be able to compute not
just grammatical sentences, but to identify which of these sentences
are \emph{good}. This problem is exacerbated when using
treebank-derived grammars, which tend to underspecify the actual
constraints on grammaticality and instead rely on statistical
information learned from the treebank.  Indeed, there have been a
number of systems for statistical generation, which can exploit such
information to rank sentences appropriately
\cite{langkildeknight1998,whitebaldridge2003,belz2008}.  However, to our 
knowledge, all such systems are currently restricted to performing surface realization,
and must rely on separate modules to perform sentence planning.

In this paper, we bring these two strands of research together for the
first time.  We present the PCRISP system, which redefines the SPUD
generation problem in terms of probabilistic TAG grammars (PTAG,
\cite{resnik1992}) and then extends CRISP to solving the probabilistic
SPUD generation problem using metric planning \cite{fox2002,hoffmann2003}.  We
evaluate PCRISP on a PTAG treebank extracted from the Wall Street
Journal Corpus \cite{chenschanker2004}.  The evaluation reveals
a tradeoff between coverage, efficiency, and accuracy which we think
are worth exploring further in future work.
